Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2024-01-02 papers/static-checker/ecoop2024/ecoop2024.tex
revise static checker

The smart contract program takes a stack as input, consisting of a single pair where the first element is an input value referred to as parameter denoted as \VPAR and the second element represents the content of the storage space denoted as \VSTORAGE. The program produces a stack as output, which is a single pair. The first element of this pair is the list of internal operations that the program intends to emit (\VOPERATIONLIST), and the second element is the new contents of the storage space.  Let's define the initial stack as  \SINIT\   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as \SFINAL\   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.
\begin{figure}[]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \NEVER \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2}\\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\ 
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}
\subsection{Rules}
Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOPLEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

@@

@@

Listing \ref{lst:smart-contract-property-specification} shows the syntax of our property specification language. Here, \KMCONTRACT, \KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KNAME\ is replaced by the contract name. In the input and output fields, the input and output patterns are specified. Since Michelson operates on a stack, we assume that the stack has only one element, as specified in the input pattern. Similarly, the output stack is assumed to be a stack with only one element, which is specified in the output field. Our syntax allows to specify incomplete and complete smart contract code. For the complete smart contrat code, as we have mentioned before, the output should be a pair consisting of an operation list and a storage.In our syntax, for verification purposes, we exclude the operation list, and the output is solely represented by the pattern of the value in the storage after the execution concludes. In fact, for a complete smart contract code, in some cases, the output pattern can be omitted, given that the input pattern must be a pair of parameters and storage. This implies that the storage pattern is part of the input pattern. In such cases, the prefix 'post' is used before the variables in the input term to indicate the corresponding storage value after the execution.  The input and output patterns represent Michelson terms, which can be either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its corresponding type declaration. However, subsequent mentions can omit the type declaration. Formally, the pre and post-condition instances are formulas in first-order logic, and they can be handled by an SMT solver, namely the Z3 solver in our framework. The detail of the language grammar is described in in Figure \ref{}. Formulas that are used in assertions or added to solvers are terms of Boolean sort. Otherwise, terms of Boolean and non-Boolean sort may be mixed in any combination where sorts match up. Universal and existential quantifiers bind variables to the scope of the quantified formula.

@@

A smart contract may offer various functions for users to interact. Therefore, our language is designed to allow users to specify different functions, referred to as entrypoints, in Michelson. This design enables us to specify and verify all functions within a smart contract. Furthermore, this design allows us to describe and verify relationships between functions.  The syntax is as follows: 
\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax},label={lst:safer-contract-module},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      code := {c2}
      input := t2
      output := t2'
      pre-condition := cd2
      post-condition := cd2'
   entrypoint %c 
      ...
   < entrypoint relations>
\end{lstlisting}
where the entrypoint name is written after the \lstinline/%/ symbol and is followed by a property specification for the entrypoint. At the end of the specification is the entrypoint relation specification, which we will explain next.

@@

\begin{lstlisting}[float,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract Auction = spec 
  entrypoint %bid
    code := {...}
    input := Pair Unit (Pair (auction_open: bool) 
                             (Pair (highest_bidder: address)
                                   (contract_owner: address)))
    output :=   Pair (post_auction_open: bool) 
                     Pair (post_highest_bidder: address)
                          (post_contract_owner: address))
    pre-condition := auction_open |
             Amount > Balance
    post-condition := post_highest_bidder = Sender |
             Post_Balance = Amount |
             Transfer_token (Unit, Balance, highest_bidder) |
             post_contract_owner = contract_owner ;
  entrypoint %close
    code := {...}
    input := Pair Unit (Pair (auction_open: bool) 
                             (Pair (highest_bidder: address)
                                   (contract_owner: address)))
    output :=   Pair (post_auction_open: bool) 
                     Pair (post_highest_bidder: address)
                          (post_contract_owner: address))
    pre-condition := auction_open |
           Sender = contract_owner                
    post-condition := not post_auction_open |
          Post_Balance = 0 |
          Transfer_token (Unit, Balance, contract_owner) ;
    (%create -> %bid) with (auction_open = true)  
                      && (Amount > Balance) | 
    (%create  -> %close) with (auction_open = true) 
                      && (Sender = contract_owner) |
    (%bid -> %bid) with (Amount > Balance)  | 
    (%bid -> %close) with (Sender = contract_owner)  | 
    not (%close -> %bid) |
    not (%close -> %close)
\end{lstlisting}

@@

\subsubsection{Property check}
The static checker enables users to verify smart contract properties by providing pre and post condition specifications. A property is specified with the following clause:  \lstinline/pre-condtion : =/ $\mid$ $\Phi_{i}$ \lstinline/ post-condtion : =/   $\mid$ $\Psi_{j}$. This specification asserts that if all preconditions $\Phi_{i}$ hold before smart contract execution, then all $\Psi_{j}$ must become true after the execution. To verify this property, we explore all the paths $\Pi$ with the path condition $\Theta$ such that ($\bigwedge \Phi_{i}$ $\land$ $\Theta$) is satisfiable. We then check that for every path $\Pi$, $\neg$ ($\bigwedge \Phi_{i} \rightarrow \bigwedge \Psi_{j}$) is unsatisfiable in the context given by the input for the entry point and its corresponding final state and path condition.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the \lstinline/transfer/ entrypoint credits the account associated with the address provided in the \lstinline/to/ parameter, while debiting the account corresponding to the \lstinline/from/ parameter with the specified \lstinline/value/ amount. The FA 1.2 standard mandates that this call must fail under the following two conditions: (1) In the event of insufficient balance, this entrypoint MUST fail with the error \lstinline/NotEnoughBalance/, indicating insufficient funds in the sender's account to execute the transfer. The error MUST contain a \lstinline/pair (nat :value) (nat :present)/ pair, where \lstinline/value/ is the requested amount of tokens, and \lstinline/present/ is the available amount. (2) If the sender's account has no permission to withdraw the \lstinline/value/ amount of funds, the call will fail with the \lstinline/NotEnoughAllowance/ error, which also contains the same pair message as the first failure scenario. To verify these requirements, we utilize the failwith condition feature of the tool and examine the returned results, as outlined in Listing \ref{}.

The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent.

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/approve/ entrypoint},label={lst:specification-approve},numbers=left]
Input := Pair (Pair (owner: address) (c: contract nat))
  (Pair (ledger: big_map address (pair nat (map address nat))) 
         (Pair (admin: address) 
               (Pair (paused: bool) (total_supply: nat))))
Pre-condition := not paused |
   get_big_map (Sender, ledger) = 
   Some (v: option (pair nat (map address nat))) | 
   v = Pair (n1: nat) (m: map address nat) |
   get_map (spender, m) = Some (n2: nat) |
   not (n2 = 0) |
   (value = 0)
Post-condition := get_big_map (Sender, post_ledger) = 
   Some (post_v: option (pair nat (map address nat))) | 
   post_v = Pair (post_n1: nat) (post_m: map address nat) |
   get_map (spender, post_m) = Some (post_n2: nat) |
   (post_n2 = 0) ;

Pre-condition := not paused |
   get_big_map (Sender, ledger) = 
   Some (v: option (pair nat (map address nat))) | 
   v = Pair (n1: nat) (m: map address nat) |
   get_map (spender, m) = Some (n2: nat) |
   (n2 = 0) |
   not (value = 0)
Post-condition := get_big_map (Sender, post_ledger) = 
   Some (post_v: option (pair nat (map address nat))) | 
   post_v = Pair (post_n1: nat) (post_m: map address nat) |
   get_map (spender, post_m) = Some (post_n2: nat) |
   (post_n2 = value) ;
\end{lstlisting}

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
Input := Pair (Pair (owner: address) (c: contract nat))
  (Pair (ledger: big_map address (pair nat (map address nat))) 
         (Pair (admin: address) 
               (Pair (paused: bool) (total_supply: nat))))
Pre-condition :=  get_big_map (owner, ledger) = 
  Some v: option (pair nat (map address nat)) |
  v  = Pair (k: nat) (m: (map address nat))
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  transfer_token (k, Amount, c) ;

Pre-condition :=  get_big_map (owner, ledger) = 
  None option (pair nat (map address nat)) 
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Other entrypoints}
The contract provides five additional entrypoints: \lstinline/mint/ to mint you own, \lstinline/burn/ to burn existing tokens, \lstinline/setAdministrator/ and \lstinline/getAdministrator/ to set up or retrieve the administrator of the smart contract, and \lstinline/setPause/ to pause the contract. To generate fail conditions, we detect that \lstinline/transfer/ or \lstinline/approve/ calls fail if the contract is in a paused state. Users or dapps should be aware of this when interacting with the smart contract.

@@

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) | 
       (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 |
       Sender = harbinger  |
       s = 'XTZ-USD' |
       Now - t <= 30 |
       client = Some a |
       get_contract (nat, a) = Some c    
    post-condition := state = 0 |
       post_client = None |
       transfer(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) |
       (post_maxDelay = maxDelay) |
       (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) |
       (post_client = client) |
       (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     && (s = 'XTZ-USD') 
     && (Now - t <= 1800)  | 
  (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  not (%getXtzUsdRate -> %getXtzUsdRate) |
  not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
\section {Conclusion}
\section {Appendix}
\end{document}

>>>>

The smart contract program takes a stack as input, consisting of a single pair where the first element is an input value referred to as parameter denoted as \VPAR and the second element represents the content of the storage space denoted as \VSTORAGE. The program produces a stack as output, which is a single pair. The first element of this pair is the list of internal operations that the program intends to emit (\VOPERATIONLIST), and the second element is the new contents of the storage space.  Let's define the initial stack as  \SINIT\   = (\KPAIR\ \VPAR\ \VSTORAGE, \TPAIR\ \TYF\ \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VPAR\ and \VSTORAGE\ denote terms representing the parameter and the storage, and \TYF\ and \TYS\ are their corresponding types. Similarly, designate the final stack as \SFINAL\   = (\PAIR\ \VOPERATIONLIST\ \VSTORAGE, \TPAIR\ (\TOPERATIONLIST) \TYS) \STACKCONCAT\ \EMPTYSTACK, where \VOPERATIONLIST\ repesents a operation list.
\begin{figure}[]
\begin{align*}
\text{t} &::= \\
   &\Mid\ \langle \text{variable} \rangle \\
   &\Mid\ \langle \text{account constant} \rangle \\
   &\Mid\ \langle \text{int constant} \rangle \\
   &\Mid\ \langle \text{string constant} \rangle \\
   &\Mid\ \langle \text{byte sequence constant} \rangle \\
   &\Mid\ \UNIT \\
   &\Mid\ \NEVER \\
   &\Mid\ \TRUE \\
   &\Mid\ \FALSE \\
   &\Mid\ \PAIR\ \text{t1 t2} \\
   &\Mid\ \LEFT\ \text{t}\\
   &\Mid\ \RIGHT\ \text{t}\\
   &\Mid\ \SOME\ \text{t}\\
   &\Mid\ \NONE \\
   &\Mid\ \text{\{t ; ... \}}\\
   &\Mid\ \text{\{ Elt t1 t2 ; ... \}}\\
   &\Mid\ \{ \langle\text{instruction}\rangle; ... \}   \\
\langle \text{variable} \rangle &::= \\ 
   &\Mid\ \VAR\\
\langle \text{account constant} \rangle &::= \\ 
   &\Mid\ \text{balance} \\
   &\Mid\ \text{amount} \\
   &\Mid\ \text{sender} \\
   &\Mid\ \text{source} \\
   &\Mid\ \text{now} \\
   &\Mid\ \text{level} \\
   &\Mid\ \text{chain-id} \\
   &\Mid\ \text{self}  \\
   &\Mid\ \text{self-address}  \\
   &\Mid\ \text{total-voting-power}  \\
   &\Mid\ \text{voting-power}  \\
\langle\text{instruction}\rangle &::= \\
  &\Mid\ \DROP \\
  &\Mid\ \DROP \langle\text{natural number constant}\rangle \\
  &\text{...}
\end{align*}
\caption{Terms}
\label{fig:term}
\end{figure}
\subsection{Rules}
Each instruction is defined by a rule on the system model. The rule semantic is defined by several kinds of transitions:
\begin{enumerate}
\item \ExprTrans\ single-step evaluation of an expression in a system state,
\item \StateTrans\ internal transitions of a state, which is non-deterministic  as shown for loop instructions, such as \LOOP\ and \LOOPLEFT,
\item \SystemTrans\ symbolic system transitions.
\end{enumerate}

@@

\begin{figure}[]
\begin{align*}
\textbf{c} & ::= 
    \text{Balance}
   \Mid \text{post\_Balance}  
   \Mid \text{Amount} 
   \Mid \text{Sender} 
   \Mid \text{Source} 
   \Mid \text{Now} 
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id} 
   \Mid \text{Self\_address} 
   \Mid \text{Total\_voting\_power} 
   \Mid \text{Voting\_power} \\
\textbf{e} & :: =  \textbf{i} \Mid  \textbf{s}  \Mid  \textbf{b} \Mid \textbf{c} \Mid (\text{x} : \text{T}) \\
&\Mid \text{Unit}  \Mid \text{Never}  \Mid \text{True}  \Mid \text{False} \\
&\Mid \text{Pair } \textbf{e } \textbf{e}  \Mid \text{Left } \textbf{e}  \Mid \text{Right } \textbf{e}  \Mid \text{Some } \textbf{e} \text{: T} \Mid \text{None T}  \\
& \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ \/ } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& \Mid \textbf{e} \text{ and } \textbf{e}\Mid \textbf{e} \text{ or } \textbf{e}  \Mid \textbf{e} \text{ nor } \textbf{e} \Mid  \text{ not } \textbf{e} \Mid \textbf{e} \text{ imply } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e}, \textbf{e}, \textbf{e}) 
 \Mid \text{get\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{get\_big\_map} (\textbf{e}, \textbf{e}) \\
& \Mid \text{get\_contract} (\textbf{e}, \textbf{e})  
 \Mid \text{abs} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{$\wedge$} \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e} 
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
&\Mid \text{forall} (\textbf{e}, \textbf{p})
\Mid \text{exsit} (\textbf{e}, \textbf{p})
\end{align*}
\caption{Syntax of predicates, p}
\label{fig:syntax-predicates}
\end{figure}

@@

Listing \ref{lst:smart-contract-property-specification} shows the syntax of our property specification language. Here, \KMCONTRACT, \KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are keywords. The \KNAME\ is replaced by the contract name. In the input and output fields, the input and output patterns are specified. Since Michelson operates on a stack, we assume that the stack has only one element, as specified in the input pattern. Similarly, the output stack is assumed to be a stack with only one element, which is specified in the output field. Our syntax allows to specify incomplete and complete smart contract code. For the complete smart contrat code, as we have mentioned before, the output should be a pair consisting of an operation list and a storage.In our syntax, for verification purposes, we exclude the operation list, and the output is solely represented by the pattern of the value in the storage after the execution concludes. In fact, for a complete smart contract code, in some cases, the output pattern can be omitted, given that the input pattern must be a pair of parameters and storage. This implies that the storage pattern is part of the input pattern. In such cases, the prefix 'post' is used before the variables in the input term to indicate the corresponding storage value after the execution.  The input and output patterns represent Michelson terms, which can be either concrete values or symbolic values, including variables. When a variable appears for the first time, it needs to be accompanied by its corresponding type declaration. However, subsequent mentions can omit the type declaration. Formally, the pre and post-condition instances are formulas in first-order logic, and they can be handled by an SMT solver, namely the Z3 solver in our framework. The specification language is constituted of predicates defined using the syntax defined as in Figure \ref{fig:syntax-predicates}. Formulas that are used in assertions or added to solvers are terms of Boolean sort. Otherwise, terms of Boolean and non-Boolean sort may be mixed in any combination where sorts match up. Universal and existential quantifiers bind variables to the scope of the quantified formula.

@@

A smart contract may offer various functions for users to interact. Therefore, our language is designed to allow users to specify different functions, referred to as entrypoints, in Michelson. This design enables us to specify and verify all functions within a smart contract. Furthermore, this design allows us to describe and verify relationships between functions.  The syntax is as follows: 
\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax},label={lst:multiple-entrypoint-specification},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      code := {c2}
      input := t2
      output := t2'
      pre-condition := cd2
      post-condition := cd2'
   entrypoint %c 
      ...
   < entrypoint relations>
\end{lstlisting}
where the entrypoint name is written after the \lstinline/%/ symbol and is followed by a property specification for the entrypoint. At the end of the specification is the entrypoint relation specification, which we will explain next.

\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax},label={lst:multiple-entrypoint-specification},numbers=left][h]
mcontract Name = spec
   storage := s
   entrypoint %a 
      code := {c1}
      parameter := t1
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      code := {c2}
      parameter := t2
      pre-condition := cd2
      post-condition := cd2'
   entrypoint %c 
      ...
   < entrypoint relations>
\end{lstlisting}

@@

\begin{lstlisting}[float,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := Pair (auction_open: bool) 
                     Pair (highest_bidder: address)
                          (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance
    post-condition := post_highest_bidder = Sender 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance, highest_bidder) 
       && post_contract_owner = contract_owner ;
  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open 
       && Sender = contract_owner                
    post-condition := not post_auction_open 
       && post_Balance = 0 
       && transfer_token (Unit, Balance, contract_owner) ;
    (%create -> %bid) with (auction_open = true)  
       && (Amount > Balance) 
    | (%create  -> %close) with (auction_open = true) 
       && (Sender = contract_owner)
    | (%bid -> %bid) with (Amount > Balance)  
    | (%bid -> %close) with (Sender = contract_owner)  
    | not (%close -> %bid) 
    | not (%close -> %close)
\end{lstlisting}

@@

\subsubsection{Property check}
The static checker enables users to verify smart contract properties by providing pre and post condition specifications. A property is specified with the following clause:  \lstinline/pre-condtion := &&/ $\Phi_{i}$ \lstinline/post-condtion := &&/ $\Psi_{j}$. This specification asserts that if all preconditions $\Phi_{i}$ hold before smart contract execution, then all $\Psi_{j}$ must become true after the execution. To verify this property, we explore all the paths $\Pi$ with the path condition $\Theta$ such that ($\bigwedge \Phi_{i}$ $\land$ $\Theta$) is satisfiable. We then check that for every path $\Pi$, $\neg$ ($\bigwedge \Phi_{i} \rightarrow \bigwedge \Psi_{j}$) is unsatisfiable in the context given by the input for the entry point and its corresponding final state and path condition.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the \lstinline/transfer/ entrypoint credits the account associated with the address provided in the \lstinline/to/ parameter, while debiting the account corresponding to the \lstinline/from/ parameter with the specified \lstinline/value/ amount. The FA 1.2 standard mandates that this call must fail under the following two conditions: (1) In the event of insufficient balance, this entrypoint MUST fail with the error \lstinline/NotEnoughBalance/, indicating insufficient funds in the sender's account to execute the transfer. (2) If the sender's account has no permission to withdraw the \lstinline/value/ amount of funds, the call will fail with the \lstinline/NotEnoughAllowance/ error, which also contains the same pair message as the first failure scenario. To verify these requirements, we utilize the failwith condition feature of the tool and examine the returned results, as outlined in Listing \ref{lst:transfer-udstz-contract}.  
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. paused: 'TokenOperationsArePaused'
2. not paused &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not paused &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not paused &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract UDStz = spec 
   storage :=  Pair (ledger: big_map address 
      (pair nat (map address nat))) 
      (Pair (admin: address) 
              (Pair (paused: bool) (total_supply: nat)))
   entrpoint %transfer
      parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
      pre-condition := not paused 
        && get_big_map (from, ledger) = 
        && Some (v: option (pair nat (map address nat))) 
        && (v = Pair (n1: nat) (m: map address nat))
        && get_map (Sender, m) = Some (n2: nat) 
        && (n1 >= value) && (n2 >= value)
      post-condition := get_big_map (from, post_ledger) = 
        && Some (post_v: option (pair nat (map address nat))) 
        && post_v = Pair (post_n1: nat) (post_m: map address nat) 
        && get_map (Sender, post_m) = Some (post_n2: nat) 
        && (post_n1 = n1 - value) | (post_n2 = n2 - value)
\end{lstlisting}

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/approve/ entrypoint},label={lst:specification-approve},numbers=left]
entrpoint %approve
   parameter := Pair (spender: address) (value: nat)
   pre-condition := not paused 
      && get_big_map (Sender, ledger) = 
        Some (v: option (pair nat (map address nat))) 
      && v = Pair (n1: nat) (m: map address nat) 
      && get_map (spender, m) = Some (n2: nat) 
      && not (n2 = 0) | (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v: option (pair nat (map address nat))) 
      && post_v = Pair (post_n1: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_n2: nat) 
      && (post_n2 = 0) ;

   pre-condition := not paused 
      && get_big_map (Sender, ledger) = 
        Some (v: option (pair nat (map address nat))) 
      && v = Pair (n1: nat) (m: map address nat) 
      && get_map (spender, m) = Some (n2: nat) 
      && (n2 = 0) | not (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v: option (pair nat (map address nat))) 
      && post_v = Pair (post_n1: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_n2: nat) 
      && (post_n2 = value) ;
\end{lstlisting}

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
entrypoint %getBalance
   parameter := Pair (owner: address) (c: contract nat)
   pre-condition :=  get_big_map (owner, ledger) = 
        Some v: option (pair nat (map address nat)) 
      && v  = Pair (k: nat) (m: (map address nat))
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = Balance 
      && transfer_token (k, Amount, c) ;

   pre-condition :=  get_big_map (owner, ledger) = 
        None option (pair nat (map address nat)) 
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = Balance 
      && transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Other entrypoints}
The contract provides five additional entrypoints: \lstinline/mint/ to mint you own, \lstinline/burn/ to burn existing tokens, \lstinline/setAdministrator/ and \lstinline/getAdministrator/ to set up or retrieve the administrator of the smart contract, and \lstinline/setPause/ to pause the contract. To generate fail conditions, we detect that \lstinline/transfer/ or \lstinline/approve/ calls fail if the contract is in a paused state. Users or dapps should be aware of this when interacting with the smart contract.

@@

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) 
       && (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 
       && Sender = harbinger  
       && s = 'XTZ-USD' 
       && Now - t <= 30 
       && client = Some a 
       && get_contract (nat, a) = Some c    
    post-condition := state = 0 
       && post_client = None 
       && transfer(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) 
       && (post_maxDelay = maxDelay) 
       && (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) 
       && (post_client = client) 
       && (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     && (s = 'XTZ-USD') 
     && (Now - t <= 1800)  
  | (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  | not (%getXtzUsdRate -> %getXtzUsdRate) 
  | not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
\section {Conclusion}
\section {Appendix}
\end{document}
